Generalized Plackett-Luce Likelihoods

### 2022-09-22 inst/jss_hyper3.Rnw
typo
fixes #156

  The \code{hyper2} package furnishes extensive idiom to work with
  extensions of the Bradley-Terry probability model such as
  Plackett-Luce likelihood functions, team strengths, and reified
  entities (monsters).  Here, I present a generalization of
  \code{hyper2} in which multiple entities are constrained to have
  identical Bradley-Terry strengths.  A new \proglang{S3} class
  \code{hyper3}, along with associated methods, is motivated and
  introduced.  Four datasets are analysed, each analysis furnishing
  new insight, and each highlighting different capabilities of the
  package.
}
\Keywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}
\Plainkeywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}

@@

\citet{mollica2014} call this a forward ranking process on the grounds
that the best (preferred; fastest; chosen) entities are identified in
the same sequence as their rank.  The \code{hyper2} package
generalizes this likelihood function to functions of ${\mathbf
  p}=(p_1,\ldots,p_n)$ with

>>>>

  The \code{hyper2} package furnishes extensive idiom to work with
  extensions of the Bradley-Terry probability model such as
  Plackett-Luce likelihood functions, team strengths, and reified
  entities (monsters).  Here, I present a generalization of
  \code{hyper2} in which multiple entities are constrained to have
  identical Bradley-Terry strengths.  A new \proglang{S3} class
  \code{hyper3}, along with associated methods, is motivated and
  introduced.  Three datasets are analysed, each analysis furnishing
  new insight, and each highlighting different capabilities of the
  package.
}
\Keywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}
\Plainkeywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}

@@

\citet{mollica2014} call this a forward ranking process on the grounds
that the best (preferred; fastest; chosen) entities are identified in
the same sequence as their rank.

Computational support for Bradley-Terry likelihood functions is
available in a range of languages.  \cite{hunter2004}, for example,
presents results in \proglang{matlab} [although he works with a
nonlinear extension to account for ties]; \cite{allison1994} present
related work for ranking statistics in \proglang{SAS} and
\cite{maystre2022} has released a \proglang{python} package for
Luce-type choice datasets.

However, the majority of software is written in the \proglang{R}
computer language~\citep{rcore2022}, which includes extensive
functionality for working with such likelihood functions:
\cite{turner2020} discuss several implementations from a computational
perspective.  The \pkg{BradleyTerry} package \citep{firth2005}
considers pairwise comparisons using \proglang{glm} but cannot deal
with ties or player-specific predictors; the \pkg{BradleyTerry2}
package \citep{turner2012} implements a flexible user interface and
wider range of models to be fitted to pairwise comparison datasets,
specifically simple random effects.  The \pkg{PlackettLuce} package
\citep{turner2020} generalizes this to likelihood functions of the
form of Equation~\ref{PL_like} and applies the Poisson transformation
of~\cite{baker1994} to express the problem as a log-linear model.  The
\pkg{hyper2} package, in contrast, gives a consistent language
interface to create and manipulate likelihood functions over the
simplex ${S}_n=\left\lbrace\left(p_1,\ldots,p_n\right)\left|p_i\geq
0,\sum p_i=1\right.\right\rbrace$.  A further extension in the package
generalizes this likelihood function to functions of ${\mathbf
p}=(p_1,\ldots,p_n)$ with
